Weather covariance matrices as U_hyp for GR50

### 2021-08-11 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Add Table 1 with summary proportions of effects with different types of GxE

# TRIAL script
# -------------------------
NB: Trial script on mash-env-cov branch to use weather covariance matrices as 
U_hyp in a mash run of GR50 phenotypes from 8 gardens.

@@

```{r load data}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
mashdir <- file.path(workingdir, "analysis", "mash")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
# pvdiv PLANT_ID metadata
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
# pvdiv common garden metadata
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
# all greenup related phenotypes, including greenup as functions of weather variables.
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
# all flowering related phenotypes, including functions of weather variables.
k_full <- read_rds(file.path(workingdir, "data",
                    "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
# kinship matrix for 630 tetraploid individuals.

@@

## 3a. Log-likelihoods
```{r}
m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
m_gulf_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_no_Uhyp.rds")
m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
m_gulf_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_no_Uhyp.rds")

m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
m_midw_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_no_Uhyp.rds")  
m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
m_midw_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_no_Uhyp.rds")

m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
m_gam_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_no_Uhyp.rds")
m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
m_gam_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_no_Uhyp.rds")
```

@@

>>>>

# Script used to make figures
# -------------------------
NB: This script uses weather covariance matrices as 
U_hyp in a mash run of GR50 phenotypes from 8 gardens.

@@

```{r load data}
source("../../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("..", "..", "..", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
mashdir <- file.path(workingdir, "analysis", "mash")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
# pvdiv PLANT_ID metadata
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
# pvdiv common garden metadata
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
# all greenup related phenotypes, including greenup as functions of weather variables.
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
# all flowering related phenotypes, including functions of weather variables.
k_full <- read_rds(file.path(workingdir, "data",
                    "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
# kinship matrix for 630 tetraploid individuals.

@@

## 3a. Log-likelihoods
```{r}
m_gulf_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
m_gulf_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_no_Uhyp.rds")
m_gulf_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
m_gulf_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_no_Uhyp.rds")

m_midw_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
m_midw_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_no_Uhyp.rds")  
m_midw_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
m_midw_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_no_Uhyp.rds")

m_gam_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
m_gam_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_no_Uhyp.rds")
m_gam_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
m_gam_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_no_Uhyp.rds")
```

@@

```{r}
site_regions <- sites %>% 
  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
                                             TRUE ~ "OK")) %>%
  select(manu_site, Region) %>%
  rename("Site" = "manu_site")

gxe_gamg$S_1_row %>%
  as_tibble(rownames = "Condition_1") %>%
  pivot_longer(cols = 2:9, names_to = "Condition_2", values_to = "Sig") %>%
  separate(Condition_1, into = c(NA, NA, NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
  separate(Condition_2, into = c(NA, NA, NA, NA, NA, "Garden_2"), sep = "_") %>%
  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
  rename("Region_1" = "Region") %>%
  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
  rename("Region_2" = "Region") %>%
  filter(Region_1 != "OK" & Region_2 !="OK") %>%
  group_by(Region_1, Region_2) %>%
  summarise(Total_sig = sum(Sig))

gxe_gf$S_AP %>%
  as_tibble(rownames = "Condition_1") %>%
  pivot_longer(cols = 2:9, names_to = "Condition_2", values_to = "Sig") %>%
  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
  rename("Region_1" = "Region") %>%
  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
  rename("Region_2" = "Region") %>%
  filter(Region_1 != "OK" & Region_2 !="OK") %>%
  group_by(Region_1, Region_2) %>%
  summarise(Total_sig = sum(Sig))

gxe_mg$S_2_no %>%
  as_tibble(rownames = "Condition_1") %>%
  pivot_longer(cols = 2:8, names_to = "Condition_2", values_to = "Sig") %>%
  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
  rename("Region_1" = "Region") %>%
  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
  rename("Region_2" = "Region") %>%
  filter(Region_1 != "OK" & Region_2 !="OK") %>%
  group_by(Region_1, Region_2) %>%
  summarise(Total_sig = sum(Sig))

gxe_mf$S_AP %>%
  as_tibble(rownames = "Condition_1") %>%
  pivot_longer(cols = 2:7, names_to = "Condition_2", values_to = "Sig") %>%
  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
  rename("Region_1" = "Region") %>%
  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
  rename("Region_2" = "Region") %>%
  filter(Region_1 != "OK" & Region_2 !="OK") %>%
  group_by(Region_1, Region_2) %>%
  summarise(Total_sig = sum(Sig))
```
